A Fast Unified Model for Parsing and Sentence Understanding

### 2016-05-31 writing/hard_stack_paper/paper.tex
Tweak shift-reduce description
Tweak introduction of 'states'.
More thin-stack tweaks

A shift-reduce parser accepts a sequence of input tokens $\mathbf x = (x_0, \dots, x_{N-1})$ and consumes transitions $\mathbf a = (a_0, \dots, a_{T-1})$, where each $a_t \in \{\shift, \reduce\}$ specifies one step of the parsing process. In general a parser may also generate these transitions on the fly as it reads the tokens. It proceeds left-to-right through a transition sequence, combining the input tokens $\mathbf x$ incrementally into a tree structure. For any binary-branching tree structure over $N$ words, this requires $T=2N-1$ transitions and $2N$ states.

@@

We propose a space-efficient stack representation inspired by the zipper technique \citep{huet1997zipper}, that we call \textit{thin stack}. For each input sentence, we represent the stack with a single $T \times D$ matrix $S$. Each row $S[t]$ (for $0<t\le T$) represents the top of the actual stack at timestep $t$. At each timestep we can \shift\ a new element onto the stack, or \reduce\ the top two elements of the stack into a single element. To shift an element from the buffer to the top of the stack at timestep $t$, we simply write it into the location $S[t]$. In order to perform the \reduce\ operation, we need to retrieve the top two elements of the actual stack. We maintain a queue $Q$ of pointers into $S$ which contains the row indices of $S$ which are still present in the actual stack. The top two elements of the stack can be found by using the final two pointers in the queue $Q$. These retrieved rows are used to perform the \reduce\ operation, which modifies $Q$ to mark that some rows of $S$ have now been replaced in the actual stack. Algorithm~\ref{alg:thin-stack} describes the full mechanics of a stack feedforward in this compressed representation. It operates on the single $T \times D$ matrix $S$ and a backpointer queue $Q$. Table~\ref{tbl:thin-stack} shows an example run.

This stack representation requires substantially less space. It stores each element involved in the feedforward computation exactly once, meaning that this representation can still support efficient backpropagation. Furthermore, all of the updates to $S$ and $Q$ can be performed batched and in-place on a GPU, yielding substantial speed gains. We describe speed results in Section~\ref{sec:speed}.

>>>>

A shift-reduce parser accepts a sequence of input tokens $\mathbf x = (x_0, \dots, x_{N-1})$ and consumes transitions $\mathbf a = (a_0, \dots, a_{T-1})$, where each $a_t \in \{\shift, \reduce\}$ specifies one step of the parsing process. In general a parser may also generate these transitions on the fly as it reads the tokens. It proceeds left-to-right through a transition sequence, combining the input tokens $\mathbf x$ incrementally into a tree structure. For any binary-branching tree structure over $N$ words, this requires $T=2N-1$ transitions through a total of $T+1$ states.

@@

We propose a space-efficient stack representation inspired by the zipper technique \citep{huet1997zipper} that we call \textit{thin stack}. For each input sentence, we represent the stack with a single $T \times D$ matrix $S$. Each row $S[t]$ (for $0<t\le T$) represents the top of the actual stack at timestep $t$. At each timestep we can \shift\ a new element onto the stack, or \reduce\ the top two elements of the stack into a single element. To shift an element from the buffer to the top of the stack at timestep $t$, we simply write it into the location $S[t]$. In order to perform the \reduce\ operation, we need to retrieve the top two elements of the actual stack. We maintain a queue $Q$ of pointers into $S$ which contains the row indices of $S$ which are still present in the actual stack. The top two elements of the stack can be found by using the final two pointers in the queue $Q$. These retrieved elements are used to perform the \reduce\ operation, which modifies $Q$ to mark that some rows of $S$ have now been replaced in the actual stack. Algorithm~\ref{alg:thin-stack} describes the full mechanics of a stack feedforward in this compressed representation. It operates on the single $T \times D$ matrix $S$ and a backpointer queue $Q$. Table~\ref{tbl:thin-stack} shows an example run.

This stack representation requires substantially less space. It stores each element involved in the feedforward computation exactly once, meaning that this representation can still support efficient backpropagation. Furthermore, all of the updates to $S$ and $Q$ can be performed batched and in-place on a GPU, yielding substantial speed gains over both a more na\"ive SPINN implementation and a standard TreeRNN implementation. We describe speed results in Section~\ref{sec:speed}.
